Estimating Real-Time Highstreet Footfall from Wi-Fi Probe Requests

### 2018-07-01 paper/sections/conclusion.tex
Final edits

This paper has made a novel contribution to the measurement of activities in retail centres in real time.
To this end, it has described the collection and processing of a novel consumer Big Dataset that enables valid measures of levels of footfall activity which has been scaled across a Great Britain wide network of sensors.
In both conceptual and technical terms, it illustrates the ways in which passively collected consumer data can be ‘hardened’ to render them robust and reliable by using related procedures of internal and external validation.

Internal validation addresses the issues of screening out device probes that do not indicate footfall, and the further screening of device probes to ‘fingerprint’ the effects of MAC randomization.It is important to note that the filtering process work based soley on the information present in the probe requests and their temporal distribution.
This ensures that although the mobile devices were uniquely identified, there was no further personal data generated by linking the probe requests to the users of the mobile devices.
This method essentially gave us a way to estimate the footfall in real-time without identifying or tracking the mobile devices themselves. 
External validation then entails reconciling adjusted counts with the footfall observed at sample locations.
This procedure makes it possible to generalise from locations at which manual footfall surveys are conducted to all others in the system, and to develop a classification of device locations that are more or less susceptible to noise generation.

This Wi-Fi based footfall counting methodology offers a large number of applications and benefits for real time spatial analysis.
Since Wi-Fi based sensors are inexpensive and the data model is scalable, it is possible to use this methodology for a large network of sensors to gather granular data on pedestrian footfall.
A snapshot showing week's worth of precise footfall in area around Charring cross, London is shown in Figure \ref{} to demonstrate the potential for such dataset.
Projects such as SmartStreetSensors \citep{sss2016}, may utilise this methodology to overcome the challenges introduced by the implementation of MAC address randomisation.
Such precise and granular data also enables us to confidently model the pedestrian flow in urban road networks, and will be an indispensable tool in the smart city framework.
It can also be used to understand and classify geographical areas based on the spatio-temporal distribution of the volume of activity in them which we intend to research in future.

>>>>

This paper has made a novel contribution to the measurement of activities in retail centres in real time.
To this end, it has described the collection and processing of a novel consumer Big Dataset that enables valid measures of levels of footfall activity which has been scaled across a Great Britain wide network of sensors \citep{sss2016}.
In both conceptual and technical terms, it illustrates the ways in which passively collected consumer data can be ‘hardened’ to render them robust and reliable by using related procedures of internal and external validation.

Internal validation addresses the issues of screening out device probes that do not indicate footfall, and the further screening of device probes to ‘fingerprint’ the effects of MAC randomization.It is important to note that the filtering process work based solely on the information present in the probe requests and their temporal distribution.
This ensures that although the mobile devices were uniquely identified, there was no further personal data generated by linking the probe requests to the users of the mobile devices.
This method essentially gave us a way to estimate the footfall in real-time without identifying or tracking the mobile devices themselves. 
External validation then entails reconciling adjusted counts with the footfall observed at sample locations.
This procedure makes it possible to generalise from locations at which manual footfall surveys are conducted to all others in the system, and to develop a classification of device locations that are more or less susceptible to noise generation.

This Wi-Fi based footfall counting methodology offers a large number of applications and benefits for real time spatial analysis.
Since Wi-Fi based sensors are inexpensive and the data model is scalable, it is possible to use this methodology for a large network of sensors to gather granular data on pedestrian footfall.
A snapshot showing week's worth of precise footfall in area around Charring cross, London is shown in Figure \ref{main_study_counts} to demonstrate the potential for such dataset.
Projects such as SmartStreetSensors \citep{sss2016}, may utilise this methodology to overcome the challenges introduced by the implementation of MAC address randomisation.
Such precise and granular data also enables us to confidently model the pedestrian flow in urban road networks, and will be an indispensable tool in the smart city framework.
It can also be used to understand and classify geographical areas based on the spatio-temporal distribution of the volume of activity in them which we intend to research in future.

### 2018-07-01 paper/sections/introduction.tex
Final edits

New and developing technologies today provide the infrastructure over which movements and interactions can be measured and monitored in the `sentient city' \citep{amin2017seeing}.
These include mobile phone networks, which can triangulate user locations relative to networks of masts, use of GPS to locate users of social media services, and WiFi beacon connectivity to access the Internet.
These technologies offer differing levels of spatial precision, where mobile telephony and WiFi generally being less reliable and offer lower precision than GPS to the end users, while being more advantageous for broader mobility studies \citep{pinelli2015comparing}.

There has been considerable research into the usefulness of these technologies to understand patterns of movement in cities in near real time \citep{candia2008uncovering, gonzalez2008understanding, calabrese2013understanding}. Most of this research has focused upon technical specification of accuracy or precision \citep{song2010limits, lane2010survey}, with somewhat less attention devoted to the ways that the characteristics of the technologies and of their human users conspire to create possible bias in the inclusiveness of usage patterns.
This is perhaps most manifest in the plethora of papers that analyse the distribution and content of publicly available Twitter social media feeds, which likely constitute an extremely biased and self-selecting subset of users that geo-enable their posts, drawn from a user base that itself bears no identified correspondence to any known background population \citep{lansley2016deriving}. 
Even analysis of mobile phone data, usually derived from industry players that have significant market share and user bases that may be representative of local populations, may exclude groups such as tourists from distand origins or subscribers to third party services that share distinctive characteristics \citep{di2016mind}

These examples illustrate the issues that underpin the assembly and analysis of consumer data, which can be considered as a distinctive class of Big Data that arise from the interactions between humans and customer-facing organisations such as retailers, domestic energy suppliers, transport providers and suppliers of social media and communications \citep{cdrc2018consumer}.
Consumer data account for an ever-increasing real share of all of the data that are collected of citizens, but a fundamental characteristic of consumer-led markets is that no single provider has a monopoly in market provision, and therefore issues of market share and segment generate bias in analysis.
The source and operation of this bias is unknown in the absence of extensive and context sensitive attempts to triangulate consumer data with data of known provenance relating to clearly defined populations \citep{lansley2016deriving}.
In similar ways to other classes of Big Data, consumer data are best throught of as digital ‘exhaust’, or a by-product created by or harvested from consumer transactions.

In this paper, through a set of experiments, we evaluate the value of data collected from a Great Britain wide network of 800 devices, installed in order to characterize the footfall patterns of a scientifically balanced sample of retail centres.
These devices are located in shop windows and record the probes emitted by mobile phones and other WiFi enabled devices.
These are deemed to be consumer data because devices carried by consumers routinely probe for a consumer service, specifically a WiFi connection.
Monitoring the probes from such devices provides an indication of the presence of their users, regardless of whether or not Internet connectivity is established.
Our core motivation is to appraise the usefulnessof WiFi probe records harvested from our network of sensors in order to indicate levels of pedestrian activity.
More broadly still, in our future research we intend to classify the nationwide network of footfall profiles as part of a programme of research to understand the form and functioning of retail areas at a time of far-reaching structural change for the retail industry.

To this end, it is important to first undertake a thorough conceptual and technical appraisal of our consumer data source.
In technical terms, screening of information present in the `probes' and classifing them based on their characteristics is essential in order to remove the ones emitted by devices that do not indicate pedestrian activity, such as network enabled printers and other fixed devices.
Related to this, a method to figerprint Wi-Fi probes is necessary to remove probes from individuals’ devices that in conceptual terms should not be considered part of footfall – as when, for example, an employees is seated in an office within range of the sensor device.
A calibration of sensor measurement is also essential on two grounds: first, individuals may carry multiple devices, or no device at all; and second, the positioning and orientation of the sensor in the retail unit may lead to systematic over- or under-enumeration.
These sources of bias in measurement must be accommodated by manual recording of footfall at each location and generalization of these sample survey results to all locations and time periods.
As we describe in detail below, manual validation of the data needs to be undertaken in parallel with technical profiling of the mix of consumer mobile devices that probe our sensors, since the effectiveness of data cleaning procedures discussed in this paper differ between individual locations and configurations.

>>>>

New and developing technologies today provide the infrastructure over which movements and interactions of people can be measured and monitored in the `sentient city' \citep{amin2017seeing}.
These include mobile phone networks, which can triangulate user locations relative to networks of masts, use of GPS to locate users of social media services, and Wi-Fi beacon connectivity to access the Internet.
These technologies offer differing levels of spatial precision, where mobile telephony and Wi-Fi generally being less reliable and offer lower precision than GPS to the end users, while being more advantageous for broader mobility studies \citep{pinelli2015comparing}.

There has been considerable research into the usefulness of these technologies to understand patterns of movement in cities in near real time \citep{candia2008uncovering, gonzalez2008understanding, calabrese2013understanding}. Most of this research has focused upon technical specification of accuracy or precision \citep{song2010limits, lane2010survey}, with somewhat less attention devoted to the ways that the characteristics of the technologies and of their human users conspire to create possible bias in the inclusiveness of usage patterns.
This is perhaps most manifest in the plethora of papers that analyse the distribution and content of publicly available Twitter social media feeds, which likely constitute an extremely biased and self-selecting subset of users that geo-enable their posts, drawn from a user base that itself bears no identified correspondence to any known background population \citep{lansley2016deriving}. 
Even analysis of mobile phone data, usually derived from industry players that have significant market share and user bases representative of local populations, may exclude groups such as tourists from distant origins or subscribers to third party services that share distinctive characteristics \citep{di2016mind}

These examples illustrate the issues that underpin the assembly and analysis of consumer data, which can be considered as a distinctive class of Big Data that arise from the interactions between humans and customer-facing organisations such as retailers, domestic energy suppliers, transport providers and suppliers of social media and communications \citep{cdrc2018consumer}.
Consumer data account for an ever-increasing real share of all of the data that are collected of citizens, but a fundamental characteristic of consumer-led markets is that no single provider has a monopoly in market provision, and therefore issues of market share and segment generate bias in analysis.
The source and operation of this bias is unknown in the absence of extensive and context sensitive attempts to triangulate consumer data with data of known provenance relating to clearly defined populations \citep{lansley2016deriving}.
In similar ways to other classes of Big Data, consumer data are best thought of as digital ‘exhaust’, or a by-product created by or harvested from consumer transactions.

In this paper, through a set of experiments, we evaluate the value of data collected from a Great Britain wide network of 800 devices \citep{sss2016}, installed in order to characterize the footfall patterns of a scientifically balanced sample of retail centres.
These devices are located in shop windows and record the probes emitted by mobile phones and other Wi-Fi enabled devices.
These are deemed to be consumer data because devices carried by consumers routinely probe for a consumer service, specifically a Wi-Fi connection.
Monitoring the probes from such devices provides an indication of the presence of their users, regardless of whether or not internet connectivity is established.
Our core motivation is to appraise the usefulness of Wi-Fi probe requests harvested from our network of sensors in order to indicate levels of pedestrian activity.
More broadly still, in our future research we intend to classify the nationwide network of footfall profiles as part of a programme of research to understand the form and functioning of retail areas at a time of far-reaching structural change for the retail industry.

To this end, it is important to first undertake a thorough conceptual and technical appraisal of our consumer data source.
In technical terms, screening the information present in the `probes requests' and classifying them based on their characteristics is essential in order to remove the ones emitted by devices that do not indicate pedestrian activity, such as network enabled printers and other fixed devices.
Related to this, a method to fingerprint Wi-Fi probes is necessary to remove probes from individuals’ devices that in conceptual terms should not be considered part of footfall – as when, for example, an employees is seated in an office within range of the sensor device.
A calibration of sensor measurement is also essential on two grounds: first, individuals may carry multiple devices, or no device at all; and second, the positioning and orientation of the sensor in the retail unit may lead to systematic over- or under-enumeration.
These sources of bias in measurement must be accommodated by manual recording of footfall at each location and generalization of these sample survey results to all locations and time periods.
As we describe in detail below, manual validation of the data needs to be undertaken in parallel with technical profiling of the mix of consumer mobile devices that probe our sensors, since the effectiveness of data cleaning procedures discussed in this paper differ between individual locations and configurations.

### 2018-07-01 paper/sections/main_study.tex
Final edits

\begin{figure}
	\begin{center}
		\includegraphics [width=0.90\linewidth] {images/main_schedule.jpeg}
		\caption{Days when the sensors were active at the corresponding location. The red square shows that manual data collection was also done.}
		\label{pilot_clustering}
	\end{center}
\end{figure}

The main study was designed to validate the results of the pilot study over different locations at different times. 
We choose five different locations across central London, to install the sensors and collect data for a long period of time. 
We also carry out manual counting on these locations along with this across different time of the day. 
We then apply the filtering based on signal strength and sequence numbers and compare with manual counts and evaluate the effectiveness of the process with the mean error per minute on these locations. 
Finally we calculate the ajustment factor for the first interval of manual counts and check if that works on the consecutive intervals.

The locations where the data were collected are shown in the table. The location are chosen for their variety of configuration and sources of noise. Location 1 is the `cleanest' of while location 2 is the one with the most complexity. The configuration, installation and data collection schedule is shown in the figure.

Though data was collected for many continous days, for the purposes of comparing with ground truth we just consider the only the data from sensors corresponding to those sessions. We have 12 sets of data over 6 different days. We have atleast two set manual counts for each location for verification of calibration.

\begin{figure}
	\centering
	\subfigure[Installation configuration of sensors at various locations]{
		\resizebox*{0.46\linewidth}{!}{\includegraphics[trim=20 6 20 6,clip]{images/main_configs.jpeg}}}\hspace{20pt}
	\subfigure[Density distribution of signal strength (lower values show higher signal strength)]{
		\resizebox*{0.46\linewidth}{!}{\includegraphics[trim=20 4 25 6,clip]{images/main_signals.jpeg}}}
	\caption{Distribution of signal strengths across locations} \label{methodology_schematic}
\end{figure}

First we see how the distribution of the signal strength varies with location and configuration. 
The density plot for signal strength is shown alogn with configuration in figure.
We can see that the signal strength distribution shows distinct patterns of high and low when the installation is that there is a clear distant source of noise but this distinction get more and more obscure as we move towards difficult installations. 
For example, location 2 is almost a normally distributed noise as it is too far to pick up any pedestrians but location 5 with a clear view of footpath and a phone shop next door shows clear distinction between the two. 
Intuitively the classfication algorithm should give us better results in the latter. 
It is important to note that we are dealing with relative singal strengths, this can vary with location and time of the analysis but we should be still be able to differentiate signal from noise. 
We run the kmeans classification algorithm and filter out the probe requests which are randomised and have signal strengths less than the second break (or the threshold). 
We then count the number of Unique MAC addresses present in every minute and remove MAC addreses which reappear within 30 minutes of previous appearance. 
We then compare this with the minute by minute aggreagation for the manual counts and find the average error per minute for the sensor count. 
The results are shown in the table.We see that the location 2 has the most error 400\% confirming our intuition while surprisingly location 3 has the least error -3\%. 
We also see that the error follows the complexity of the installation. 
It is also very promising that this method alone reduces our margin of error by a 50 - 100\%
For some practical purposes which do not require absolute numbers, this should be sufficient. 
e.g. Indexes of activity and change dashboards and short term trends identification.

We see that the success of the signal strength filtering depends on the how we tackle the problems in the installing the devices.
The more the device in installed in a way that the field of measurement is distinct from its surroundings (however noisy they might be) the more successful this process is.
Closer sources error causes overcounting as we see in 2, 4 but at the same time highly close field of measurement with no big source of error can cause undercounting such as in 3. The complexity of the installation and uniformity of surroundings near the sensors directly coorrelate to the MAPE.
The most effective way of measuring pedestrian is to have the sensor next to have the sensor next to clear area of pedestrian activity and in case of absence of any major source of noise, we should not do the signal strength filtering at all. We can even note the major possible sources of noise around the device and use it to fiugre out the right way of partitioning the data e.g. location 2 and 3 should have different pattern of signal strength filtering depending on their configuration.

In terms of filtering by sequence numbers we see that the the filtering works in all the locations. It has the similar effect as the signal strength and brings us furthe closer to the manual count. Finally we calculate the adjustment factor from the first 15 minutes and apply that for the rest of the interval. We get the counts down to almost 10\% on good locations. The locations which still over count are the ones with seating next to them (2 and 4). We also see that the trends are preserved with the filtering process.

\begin{table}
	\tbl{Data collected at different locations and the correspondin mean absolute percentage error after each filtering process}
	{\begin{tabular}{ccccccc} 
		\toprule
			Sensor & Signal strength & Adjustment & MAPE & MAPE after & MAPE after & MAPE of\\
			& threshold & factor & without any & filtering signal & filtering sequence & final adjusted\\
			& (-dBm) & &  cleaning (\%) & strength (\%) & numbers (\%) & counts (\%)\\
		 \midrule
			1 & -70 & 1.25 & 259 &  22 & -13 &  9 \\
			2 & -74 & 0.51 & 928 & 396 & 206 & 55 \\
			3 & -72 & 1.60 &  87 & -19 & -31 & 10 \\
			4 & -70 & 0.88 & 498 & 142 &  52 & 33 \\
			5 & -72 & 0.80 & 473 &  84 &  38 & 11 \\
		 \bottomrule
	\end{tabular}}
	\label{errors-table}
\end{table}

\begin{figure}
	\begin{center}
		\includegraphics [width=\linewidth,trim=6 6 6 6,clip] {images/main_comparison.jpeg}
		\caption{Comparison of all the counts in teh sensors}
		\label{main_comparison}
	\end{center}
\end{figure}

>>>>

\begin{figure}
	\begin{center}
		\includegraphics [width=0.90\linewidth] {images/main_schedule.jpeg}
		\caption{Data collection schedule showing the days when sensors were active at their corresponding locations. The red squares show that manual counting of pedestrians was also done on that day.}
		\label{main_schedule}
	\end{center}
\end{figure}

The main study was designed to validate the results of the pilot study over different locations at different times. We start by choosing five different locations across central London, install the sensors and collect data for extended periods of time. 
We also carry out manual counting at these locations across different times of the day. 
We then apply the methodologies discussed earlier to arrive at estimated pedestrian footfall and compare them with corresponding manual counts.
We finally evaluate the effectiveness of the processes with the Mean Absolute Percentage Error (MAPE) at these locations and report our findings.

The locations at which the data were collected are shown in Table \ref{locations-table}. The locations are chosen for their diverse site conditions and unique sources of noise around the potential location of the sensors. The position of the sensor at these locations with respect to the context is shown the Figure \ref{main_signal_strength}. We can see that Location 5 is the `cleanest' with one clear stationary source of noise (phone shop) of while location 2 is the one with the most complexity with seating area all around the sensor. 
The sensors were operational through out February and March while manual counts were conducted in these location in half hour sessions on at least two different days. For the purposes of comparing with ground truth we just considered the data from sensors which corresponding to the 12 sets of available manual counts. The schedule of data collection is shown in Figure \ref{main_schedule}

\begin{figure}
	\centering
	\subfigure[Installation configuration of sensors at the survey locations (not to scale)]{
		\resizebox*{0.46\linewidth}{!}{\includegraphics[trim=20 6 20 6,clip]{images/main_configs.jpeg}}}\hspace{20pt}
	\subfigure[Density distribution of signal strength reported in collected probe requests (lower values show higher signal strength)]{
		\resizebox*{0.46\linewidth}{!}{\includegraphics[trim=20 4 25 6,clip]{images/main_signals.jpeg}}}
	\caption{Distribution of signal strengths across locations} \label{main_signal_strength}
\end{figure}

We start by looking at the distribution of the signal strength reported by the probe requests across the locations. From the density plot shown in Figure \ref{main_signal_strength} we can see that there is clear relation between the distribution of the signal strength and the distance and complexity of the source of noise. We can see that while location 5 shows clean difference between low and high signal strengths, location 2 is almost normally distributed.
Intuitively we expect that location 2 and 4 must be harder to classify than location 1 3 and 5. 
We run the k-means classification algorithm and filter out the probe requests which are randomised and have signal strengths less than the second break (threshold). 
It is important to note that we are dealing with relative thresholds of signal strengths which can vary with location and time of the analysis.
We then aggregate then probe requests for every minute by counting the number of Unique MAC addresses present in every minute. We also remove devices that dwell around the sensor by removing the MAC addresses which reappear from within the past hour.
The results of the first stage of filtering process along with the thresholds are shown in Table \ref{errors-table}. Confirming our intuition, we see that the location 2 has the most MAPE followed by location 4 while rest of them have highly reduced MAPE.
It is significant that this method alone reduces our margin of error by a 50 - 100\% from the raw counts without any cleaning.
This makes the signal strength filtering quick and ideal method for practical applications which doesn't require absolute numbers such as creating large aggregated indexes to show long term trends. 
We also see that the success of the signal strength filtering can be improved significantly by installing sensors such that the pedestrians and noise are at different distances from the sensor and are the field of measurement is distinct from the surroundings however noisy it might be.

In terms of filtering by sequence numbers we see that the filtering works in all the locations. It has the similar effect as the signal strength and brings us further closer to the manual count. Finally we calculate the adjustment factor from the first 15 minutes and apply that for the rest of the interval. We get the counts down to almost 10\% on good locations. The locations which still over count are the ones with seating next to them (2 and 4). We also see that the trends are preserved with the filtering process.

\begin{table}
	\tbl{Data collected at different locations and the corresponding mean absolute percentage error after each filtering process}
	{\begin{tabular}{ccccccc} 
		\toprule
			Sensor & Signal strength & Adjustment & MAPE & MAPE after & MAPE after & MAPE of\\
			& threshold & factor & without any & filtering signal & filtering sequence & final adjusted\\
			& (-dBm) & &  cleaning (\%) & strength (\%) & numbers (\%) & counts (\%)\\
		 \midrule
			1 & -70 & 1.25 & 259 &  22 & -13 &  9 \\
			2 & -74 & 0.51 & 928 & 396 & 206 & 55 \\
			3 & -72 & 1.60 &  87 & -19 & -31 & 10 \\
			4 & -70 & 0.88 & 498 & 142 &  52 & 33 \\
			5 & -72 & 0.80 & 473 &  84 &  38 & 11 \\
		 \bottomrule
	\end{tabular}}
	\label{errors-table}
\end{table}

\begin{figure}
	\begin{center}
		\includegraphics [width=\linewidth,trim=6 6 6 6,clip] {images/main_comparison.jpeg}
		\caption{Comparison of all the counts in the sensors}
		\label{main_comparison}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics [width=\linewidth] {images/main_schedule.jpeg}
		\caption{A week of pedestrian footfall at Camden Highstreet collected by the methodology. The counts are aggregated for 5 minute intervals.}
		\label{main_study_counts}
	\end{center}
\end{figure}

### 2018-07-01 paper/sections/previous_work.tex
Final edits

Probe requests are low level signals standardised by IEEE 802.1b/g specification \citep{ieee2013} as the first step in establishing a Wi-Fi based connection between two devices and is implemented in any Wi-Fi capable device irrespective of the manufacturer or the model.
This ubiquity and standardisation make them an excellent source of open, passive, continuous, and wireless data generated by Wi-Fi capable devices present at any given time and location.
Considering the unprecendented levels of mobile device ownership in recent years, we can in turn use this data to understand the population distribution in highly dynamic urban environments with high spatial and temporal granularity \citep{freud2015,konto2017}. While a Wi-Fi based method to collect data offers us various advantages such as, easy scalability and efficiency in terms of cost and time, It also introduces few systematic biases, uncertainities in the collected data along with the serious risk of infringing on the privacy of the mobile users.
In this paper, using a set of probe requests and manual counts collected at various high street locations across London, we demonstrate that pedestrian footfall at these locations can be estimated with considerable precision and accuracy while protecting the privacy of the pedestrians.

Though Wi-Fi is a `location-less' technology, there are reliable methods to triangulate the location of Wi-Fi enabled mobile devices by the known locations of APs and the signal strength reported by them from the mobile devices \citep{he2003range, moore2004robust, lamarca2005place}.
This can overcome the usual shortcoming of GPS, which struggles for precision and accuracy in indoor and densely built environments \citep{zarim2006,kawaguchi2009wifi, xi2010locating}. 
Utilising this, we can easily and quickly estimate trajectories of the mobile devices just using the Wi-Fi communication the device has with multiple known APs \citep{Sorensenlocation} which can be used similar to the GPS trajectories to understand individual travel patterns (Kim, 2006;\citep{reki2007,Sap2015}, crowd behaviour \citep{abedi2013bluetooth,mowafi2013tracking}, vehicular \citep{lu2010vehicle} and pedestrian movement \citep{xu2013pedestrian,fukuzaki2014pedestrian,wang2016gait}.
Such data can also be used in transportation planning and management to estimate travel time \citep{musa2011wiflow} and real time traffic monitoring \citep{abbott2013empirical}.

Using techniques demonstrated by \cite{franklin2006passive} and \cite{pang2007802} and using the information present in the probe requests one can also successfully track people across the sensors or access points collecting them \citep{cunche2014linking}, infer their trajectories \citep{musa2012tracking}, and even model interactions between them \citep{cheng2012inferring,barbera2013signals,cunche2014know} such as predicting which of them are most likely to meet again \citep{cunche2012know}.
Using the semantic information present in these probe requests it even is possible to understand the nature of population at a large scale \citep{di2016mind}.

Though extensive research has been carried out on this subject with feasible and favorable results, in recent years, one of the major challenges faced in such attempts has been the increasing attempt by mobile phone manufacturers to protect the users’ privacy by anonymising the globally identifiable portion of the probe requests, \citep{green2008}.
There are various methods which have been devised to overcome this anonymisation process such as decomposition of OUIs where detailed device model information is estimated by analysing an already known dataset of OUIs \citep{martin2016decomposition}; Scrambler attack using a small part of the physical layer specification for Wi-Fi \citep{vo2016,bloessl2015scrambler}; and finally, the timing attack where the packet sequence information present in the probe request frame is used \citep{matte2016,cheng2016can}.
A combination of these methodologies has been proven to produce de-anonymised globally unique device information \citep{vanhoef2016, martin2017}.
These approaches usually result in serious risk of infringement of the privacy of the users of the mobile devices by revealing their identifiable personal information.

There is a clear gap in the research for exploring methodologies which enable us to estimate the number of unique mobile devices from a set of anonymised probe requests, without the need to reveal their original MAC addresses.
Such technique has various applications in numerous fields such as uncovering the urban wireless landscape \citep{rose2010mapping}, revealing human activity at large scales \citep{qin2013discovering}, estimating pedestrian numbers in crowds \citep{schauer2014estimating,fukuzaki2015statistical} and even counting people in hyper local scales such as queues \citep{wang2013measuring}
With enough infrastructure to collect such information we can aim to generate a real-time census of the city \citep{konto2017}.
It has also been demonstrated by \citep{pinelli2015comparing} through series of experiments on a telecom operator dataset that using such network-driven approach is more advantageous compared to the widely used event-driven approaches.
With this background we set out to device and implement a methodology to reliably estimate human activity such as pedestrian footfall from Wi-Fi probe requests without infringing the privacy of the users involved.

>>>>

Probe requests are low level signals standardised by IEEE 802.1b/g specification \citep{ieee2013} as the first step in establishing a Wi-Fi based connection between two devices and is implemented in any Wi-Fi capable device irrespective of the manufacturer or the model.
This ubiquity and standardisation makes them an excellent source of open, passive, continuous, and wireless data generated by Wi-Fi capable devices present at any given time and location.
Considering the unprecendented levels of mobile device ownership in recent years, we can, in turn use this data to understand the population distribution in highly dynamic urban environments with high spatial and temporal granularity \citep{freud2015,konto2017}. While a Wi-Fi based method to collect data offers us various advantages such as, easy scalability and efficiency in terms of cost and time, It also introduces few systematic biases, uncertainities in the collected data along with the serious risk of infringing on the privacy of the mobile users.
In this paper, using a set of probe requests and manual counts collected at various high street locations across London, we demonstrate that pedestrian footfall at these locations can be estimated with considerable precision and accuracy while protecting the privacy of the pedestrians.

Though Wi-Fi is a `location-less' technology, there are reliable methods to triangulate the location of Wi-Fi enabled mobile devices from the locations of known access points(AP) and the signal strength reported by them \citep{he2003range, moore2004robust, lamarca2005place}.
This can overcome the usual shortcoming of GPS, which struggles for precision and accuracy in indoor and densely built environments \citep{zarim2006,kawaguchi2009wifi, xi2010locating}. 
Utilising this, we can easily and quickly estimate trajectories of the mobile devices \citep{Sorensenlocation,musa2012tracking} which can be used similar to the GPS trajectories to understand individual travel patterns \citep{reki2007,Sap2015}, crowd behaviour \citep{abedi2013bluetooth,mowafi2013tracking}, vehicular \citep{lu2010vehicle} and pedestrian movement \citep{xu2013pedestrian,fukuzaki2014pedestrian,wang2016gait}.
Such data can also be used in transportation planning and management to estimate travel time \citep{musa2011wiflow} and real time traffic monitoring \citep{abbott2013empirical}.
Using techniques demonstrated by \cite{franklin2006passive} and \cite{pang2007802} along with information present in the probe requests one even model interactions between the users \citep{cheng2012inferring,barbera2013signals,cunche2014know,cunche2014linking} such as predicting which of them are most likely to meet again \citep{cunche2012know}.
Using the semantic information present in these probe requests it even is possible to understand the nature of population at a large scale \citep{di2016mind}.

Though extensive research has been carried out on this subject with feasible and favorable results, in recent years, one of the major challenges faced in such attempts has been the increasing attempt by mobile phone manufacturers to protect their users’ privacy by anonymising the globally identifiable portion of the probe requests, \citep{green2008}.
Various methods have been devised to overcome this anonymisation process such as estimating the device model information from a known dataset of manufacturers and device behaviours \citep{martin2016decomposition}; Scrambler attack using a small part of the physical layer specification for Wi-Fi \citep{vo2016,bloessl2015scrambler}; and timing attack where the packet sequence information along with information elements present in the probe request frame is used \citep{matte2016,cheng2016can}.
A combination of these methodologies has been proven to produce de-anonymised globally unique device information \citep{vanhoef2016, martin2017}.
These approaches usually result in serious risk of infringement of the privacy of the users of the mobile devices by revealing their identifiable personal information.

There is a clear gap in research for exploring methodologies for estimating the number of unique mobile devices from a set of anonymised probe requests, without the need to reveal their original device information.
Such technique has various applications such as uncovering the urban wireless landscape \citep{rose2010mapping}, revealing human activity at large scales \citep{qin2013discovering}, estimating pedestrian numbers in crowds \citep{schauer2014estimating,fukuzaki2015statistical} and even counting people in hyper local scales such as queues \citep{wang2013measuring}
With enough infrastructure to collect such information we can even aim to generate a real-time census of the city \citep{konto2017}.
With this background we set out to device and implement a methodology to reliably estimate human activity such as pedestrian footfall from Wi-Fi probe requests without infringing the privacy of the users involved.
